The effects of host galaxy properties on merging compact binaries detectable by LIGO

### 2016-07-31 paper/mnras/paper.tex
brooks comments

\author[R. O'Shaughnessy et al.]{R. O'Shaughnessy,$^1$\thanks{E-mail: oshaughn@mail.rit.edu}
J. M. Bellovary,$^2$ 
C. R. Christensen,$^3$ 
S. Shen,$^4$ 
F. Governato,$^5$ 
A. Brooks$^6$\\
%
%
%% List of institutions
$^1$Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA  \\
$^2$Department of Physics, Queensborough Community College, Bayside, NY 11364, USA\\
$^3$Department of Physics, Grinnell College, Grinnell, IA, 50112, USA\\
$^4$Cambridge or Oslo?\\
$^5$Department of Astronomy, University of Washington, Seattle, WA 98195, USA\\
$^6$Rutgers, New Jersey
 }

@@

%\email{fabio@astro.washington.edu}
\begin{abstract}
% POINT:
% - existence proof by example .. not an attempt to solve the general problem we raise
Cosmological simulations of galaxy formation can produce present-day galaxies with a large range assembly and star formation histories.  A detailed study of the metallicity evolution and star formation history of such simulations can assist in predicting LIGO-detectable compact object binary mergers.
Recent simulations of compact binary evolution suggest the compact object merger rate depends
sensitively on the progenitor's metallicity.  Rare low-metallicity star formation during galaxy
assembly can produce more detected compact binaries than typical star formation.   
Using detailed simulations of galaxy and chemical evolution, we determine how sensitively the compact binary populations
of galaxies   with similar
present-day appearance depend on the details of their assembly.
We also demonstrate by concrete example the extent to which dwarf galaxies overabundantly produce compact binary
mergers, particularly binary black holes, relative to more massive galaxies. 
We discuss the implications for transient multimessenger astronomy with compact binary sources.

@@

% POINT: Huge detail possible
The host galaxies of distant short GRBs have already been extensively investigated, with the associations being used to
draw preliminary conclusions about their progenitors  \citep{2014ARAA..52...43B}.   
%Of necessity, these analyses  on bulk correlations, often stacked to increase significance across multiple sources
%\editremark{clean up/don't piss off   Edo}.  
Unlike GRBs, detected gravitational wave sources will be limited by the range of LIGO to the local universe; for
example,  binary neutron star sources should be closer than $400\unit{Mpc}$.  Due to their proximity, each host galaxy
can be explored at great depth and detail via  position-resolved spectroscopy, enabling detailed position-resolved star-formation
and chemical evolution histories  \citep[see,e.g.][]{2009MNRAS.396..462K,2014MNRAS.444..336C,
CALIFA,CALIFA2}.  
%
However, unlike short GRBs and supernovae (SN), present-day compact  binary populations can depend sensitively on rare low metallicity
star formation.
% POINT: But is it useful?
In this work, we assess by concrete example the extent to which detailed analysis of individual galaxies' assembly histories will be essential in
investigations key physical questions about the origin of compact binary mergers.

This paper is organized as follows. 
In \S \ref{sec:sims}, we describe detailed hydrodynamical simulations of  several galaxies, including four of Milky Way-mass and two dwarfs.  Though the four Milky Way-like galaxies are morphologically similar at $z = 0$, we
show that their star formation and chemical evolution history have subtle but important differences due to their
distinctive merger histories.   To demonstrate
these differences have a practical impact,  in \S \ref{sec:model}, we introduce a simple,
metallicity-dependent  phenomenological model to calculate the present-day rate and mass distribution of compact binary
mergers from  a galaxy's known history.   In \S \ref{sec:results:BBH}, we use this model to investigate  the compact
binary  coalescence rate
dependence on the galaxy's assembly history.   
%% We also demonstrate that even the coalescence rates of binary neutron stars (NSNS) and black hole-neutron
%% star (BH-NS) binaries also depend surprisingly sensitively on the assembly history of each galaxy. 
In \S~\ref{sec:Discussion} we discuss the implications of our study for the interpretation of host galaxy associations
identified via transient multmiessenger astronomy, in the short and long term.  We summarize our results in \S~ \ref{sec:conclude}.
%% In \S \ref{sec:results} we show that while these two galaxy formation models should produce comparable detectable binary neutron star
%% populations, these two scenarios could produce dramatically different populations of electromagnetically detectable
%% merging black hole-neutron star binaries. 
%

@@

\subsection{Simulating galaxy evolution}
To thoroughly study the significance of low metallicity star
formation, we examine cosmological smoothed particle hydrodynamics
(SPH) $N$-body simulations of Milky Way-like galaxies with GASOLINE
\citep{Stadel01,Wadsley04}.  These simulations allow us to analyze both
spatially and temporally resolved star formation, and determine not
only the metallicity history of compact object progenitors.%, but also their dynamical evolution.  
%maybe move the below section to later and/or consolidate
%One simulation we report here, h258,
%involves the formation of a Milky Way-like disc galaxy whose
%progenitors undergo major mergers at $z = 2$ and $z = 0.8$.  This
%simulation has been previously discussed by
%\citet{Governato09,Bellovary10,Bellovary11,Brooks11}; however we have
%re-run the simulation at a factor of two higher in spatial resolution
%and additionally included new physics including a prescription for
%cooling by molecular hydrogen \citet{Christensen12}.

We selected our simulated regions of interest from a volume of
uniform resolution, and resampled the region at very high
resolution using the volume renormalization technique \citep{Katz93}.
This technique allows us to follow the detailed physical processes
involved in galaxy evolution in our selected region while still
including large-scale torques from cosmic structure.  Our cosmological parameters are $\Omega_0 = 0.24$, $\Omega_{\rm baryon} = 0.04$, $\Lambda = 0.76$, h = 0.73, $\sigma_8 = 0.77$ 
 \citep{WMAP3} and model the ionizing UV background
with the prescription from \citet{Haardt96}.  Our interstellar medium (ISM) model includes the non-equilibrium formation and destruction of H$_2$, which is incorporated in the cooling model along with metal lines \citep{Christensen12}.  Stars form
probabilistically from gas particles which meet density ($n_{min} =
0.1$ amu cm$^{-3}$) and temperature ($T_{max} = 10^3$ K) thresholds. If a gas
particle meets these criteria, it has a likelihood of forming a star
particle (representing a simple stellar population with a Kroupa IMF
\citep{Kroupa}) which is given by

@@

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{Kennicutt89};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses; \footnote{Gas particles start with a set mass
 and may gain mass from feedback and lose it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{Christensen10} for a discussion
of resolution issues in SPH simulations.} 
 $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  A detailed study of different interstellar medium (ISM) models and the resulting star formation properties by \citet{Christensen14a} demonstrates that this model allows star formation to occur in dense clumps of gas, in regions that correspond best to star formation in observed galaxies.

We model supernova feedback using the
blastwave formalism described in \citet{McKee77} and implemented in
our simulations as in \citet{Stinson06}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{Brooks07}, % JB: removed Kirby citation
 the Tully-Fisher relation \citep{Governato09}, and
the size-luminosity relation \citep{Brooks11}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{Governato10} and the Milky Way \citep{Guedes11}.

@@

We identify individual galaxies using the halo finder $AHF$
\citep{Gill04,Knollmann09}, which identifies haloes based on an
overdensity criterion for a flat universe \citep{Gross97}.  In each simulation, we are focusing on the stars which make up the primary (i.e. most massive) galaxy at $z = 0$.%, which serves as the progenitor to our $z = 0$ Milky
%Way-like galaxy.

@@

%{\bf Richard go ahead and edit this however is needed.}  
The evolution
of a galaxy, in terms of its stellar mass and metallicity evolution,
depends strongly on its interaction history.  Galaxies which appear
similar at the present day may have had drastically different
histories, which may result in differences in compact object merger
rates.  To investigate whether galaxy history affects the compact binary event
rate, we have chosen four simulations which are morphologically similar at $z
= 0$ but differ strongly in their merger histories; see Figure \ref{fig:images}.

The simulation h277  is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{Loebman12,Loebman14,Kassin14}, baryon fraction
\citep{Munshi13}, halo properties \citep{Zolotov09,Zolotov10}, and
disc structure \citep{Brooks11}.  It has a virial mass of $M_{vir} =
6.79 \times 10^{11} \msun$, stellar mass $M_* = 4.24 \times 10^{10}
\msun$, and maximum circular velocity $v_{\rm circ} = 235$ km s$^{-1}$.

@@

We include two additional Milky Way simulations with similar $z = 0$ properties, which have evolutionary histories that fall in between the extremes of the two described above.  The galaxy h239 has a total virial mass of $M_{vir} = 9.3 \times 10^{11}\msun$, stellar mass $M_* = 4.50 \times 10^{10}
\msun$, and maximum circular velocity $v_{\rm circ} = 250$ km s$^{-1}$.  The galaxy h285 has a total virial mass of $M_{vir} = 8.82 \times 10^{11}\msun$, stellar mass $M_* = 4.56 \times 10^{10} \msun$, and maximum circular velocity $v_{\rm circ} = 248$ km s$^{-1}$.  These galaxies are also described in \citet{Munshi13,Bellovary14}.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/boring}
\includegraphics[width=\columnwidth]{Figures/exciting}
\caption{\label{fig:GalaxyImages}Synthetic SDSS $gri$ images of two of our Milky Way-type galaxies, h277 (left) and h258 (right), created with \textsc{SUNRISE} \citep{Jonsson06}.
\label{fig:images}
}
\end{figure*}

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/sfh}
\includegraphics[width=\columnwidth]{Figures/zvst}
\caption{\label{fig:TwoGalaxies}\textbf{Star formation and metallicity versus time}:  \emph{Left panel}: Star formation history $\dot{M}_{*}$
  versus time.  Black corresponds to \BoringGalaxy{} and red to \ExcitingGalaxy.  
\emph{Right panel}: A plot of the  metallicity $Z$ of recently-formed stars versus time.  Solid red and black lines show
the mean metallicity; dotted lines correspond to  90\% of newly-born stars have lower metallicity,   or 10\% of
newly-born stars have greater metallicity.  When these two galaxies are $2-3\unit{Gyr}$ old, prior to the first major
merger of  near the peak of their
star-forming history, newly-born stars are created with significantly different metallicity.  Additionally, prior to the
second major merger of galaxy B at $\simeq 6\unit{Gyr}$, stars form in the \ExcitingGalaxy{} galaxy at a systematically lower
metallicity than in the \BoringGalaxy{} counterpart.
}
\end{figure*}

@@

We have also employed the results of detailed simulations for two dwarf galaxies: \DwarfOne{} and \DwarfTwo{}.  The simulation of h603 consists of a low-mass disc galaxy (qualitatively similar to M33).  It has a virial mass of $3.4 \times 10^{11} \msun$, stellar mass of $7.8 \times 10^9 \msun$, and maximum circular velocity of 111 km s$^{-1}$.  The structure and star formation of this galaxy has been extensively studied by \citet{Christensen14b}.  We also include a bulgeless dwarf galaxy, h516, which has a disc with irregularly distributions star formation, with a virial mass of $3.8 \times 10^{10}\msun$, stellar mass $2.5 \times 10^8 \msun$, and maximum circular velocity of 65 km s$^{-1}$.  Images of these galaxies are shown in Figure \ref{fig:dwarfimages}, and we show their star formation history and metallitity evolution in  Figure \ref{fig:TwoDwarfGalaxies}.  Note that these galaxies have quite different masses, and are not meant to be directly comparable.  The more massive h603 has a much more active star formation history and an overall increasing metallicity with time, whereas the less massive h516 is characterized by small bursts of star formation and a fairly flat metallicity evolution.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/dwarf1}
\includegraphics[width=\columnwidth]{Figures/h516}
\caption{\label{fig:GalaxyImages}Synthetic SDSS $gri$ images of our two lower-mass galaxies, h603 (left) and h516, created with \textsc{SUNRISE} \citep{Jonsson06}.
\label{fig:dwarfimages}
}
\end{figure*}

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/sfh_dwarf}
\includegraphics[width=\columnwidth]{Figures/dwarf_zvst}
\caption{\label{fig:TwoDwarfGalaxies}\textbf{Star formation and metallicity versus time}:  \emph{Left panel}: Star formation history $\dot{M}_{*}$
  versus time.  Black corresponds to h603 and red to h516.  
\emph{Right panel}: A plot of the  metallicity $Z$ of recently-formed stars versus time.  Solid red and black lines show
the mean metallicity; dotted lines correspond to  90\% of newly-born stars have lower metallicity,   or 10\% of
newly-born stars have greater metallicity.  %When these two galaxies are $2-3\unit{Gyr}$ old, prior to the first major
%merger of  near the peak of their
%star-forming history, newly-born stars are created with significantly different metallicity.  Additionally, prior to the
%second major merger of galaxy B at $\simeq 6\unit{Gyr}$, stars form in the \ExcitingGalaxy{} galaxy at a systematically lower
%metallicity than in the \BoringGalaxy{} counterpart.
}
\end{figure*}

@@

\abbrvPSgrb{} and \abbrvPSellipticals{} show this distribution is reasonable approximations to compact binary delay

@@

The close agreement between the two galaxies' binary black hole populations, determined by the anticorrelation between
star formation rate and metallicity, may be a single example of a broad trend.  This anticorrelation could cause
galaxies with similar present-day properties to always have similar present-day binary black hole populations,
regardless of their detailed assembly histories.
However, it is also possible that this close agreement is entirely a
a numerical coincdence between the specific star formation rate and metallicities in these two galaxies.  The effects described here
  are well within the scatter around the mass-metallicity relation.  Marginally different realizations of these
  histories can easily produce factors of order unity difference in galaxies with otherwise indistinguishable
  present-day properties.  That said, two other Milky-Way-like  galaxies (h239 and h285) also show very
  similar binary black hole coalescence rates $r_D/M$; see Table \ref{tab:Results}. % \editremark{but: do more examples show dramatic differences, or comparable changes}

@@

\begin{table}
\begin{centering}
\begin{tabular}{llllc}\hline
\text{Simulation}  &  \text{BHBH} &  \text{BHNS}  & \text{NSNS} & M$_* (\msun)$\\
\hline
\hline
 \text{\BoringGalaxy} & 0.0224 & 0.000247 & 0.000206  & $4.24 \times 10^{10}$ \\
 \text{\ExcitingGalaxy}  &0.0216 & 0.000297 & 0.000228 & $4.46 \times 10^{10}$ \\
  \text{h239}  &  0.023 & 0.000354 & 0.000268 & $4.50 \times 10^{10}$\\
 \text{h285} &  0.0236 & 0.000284 & 0.000241 & $4.56 \times 10^{10}$ \\ \hline
 \text{\DwarfOne} &  0.0381 & 0.000308 & 0.000441& $7.8 \times 10^9$ \\
 \text{\DwarfTwo}  & 0.0949 & 0.000257 & 0.000884& $2.5 \times 10^8$ \\
\hline
\end{tabular}
\end{centering}
\caption{\label{tab:Results}\textbf{Event rates per unit mass $r_D/M$, arbitrary units}, for the three different types of binaries
  discussed, along with the stellar mass of each galaxy.  
}
\end{table}

@@

While a gravitational wave detection provides detailed information about merging objects (i.e. masses, spins, distance), we need further knowledge to understand the actual origins of the progenitors.  Even when electromagnetic counterparts are available, the long delay times between binary formation and merger limit the prospects of examining the environment where the binary first formed.  On the one hand, compact binaries are kicked by supernovae, moving substantially away from their birth position
\citep{2013ApJ...776...18F,2014ARAA..52...43B}. 
On the other hand, particularly during the early epoch of galaxy formation, galaxies are well-mixed: stars and adjacent gas generally do not have similar chemical composition.  These mixing effects have been previously recognized as an obstacle to interpreting transient event spectra.  For
example,   \citet{2010MNRAS.402.1523P} previously demonstrated that absorbing gas neighbouring transient events (there, long GRBs)
would generally have high metallicity, even for low-metallicity progenitors.   
%
\citet{2010MNRAS.402.1523P} have previously used hydrodynamical simulations to demonstrate that observed ambient
metallicities (there, using damped Lyman $\alpha$ absorbers in the host) do not tightly constrain the metallicity
distribution of the progenitor; see, e.g., their Fig 3.  Thus, the metallicity of stars and gas adjacent to a specific merger event provides
few direct, unambiguous clues to a compact binary merger's progenitors.

>>>>

\author[R. O'Shaughnessy et al.]{R. O'Shaughnessy,$^1$\thanks{E-mail: oshaughn@mail.rit.edu}
J. M. Bellovary,$^2$ 
C. R. Christensen,$^3$ 
S. Shen,$^4$ 
F. Governato,$^5$ 
A. Brooks$^6$\\
%
%
%% List of institutions
$^1$Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA  \\
$^2$Department of Physics, Queensborough Community College, Bayside, NY 11364, USA\\
$^3$Department of Physics, Grinnell College, Grinnell, IA, 50112, USA\\
$^4$Institute of Astronomy, University of Cambridge, Cambridge CB3 OHA, UK\\
$^5$Department of Astronomy, University of Washington, Seattle, WA 98195, USA\\
$^6$Department of Physics and Astronomy, Rutgers University, Piscataway, New Jersey, 08854, USA
 }

@@

%\email{fabio@astro.washington.edu}
\begin{abstract}
% POINT:
% - existence proof by example .. not an attempt to solve the general problem we raise
Cosmological simulations of galaxy formation can produce present-day galaxies with a large range of assembly and star formation histories.  A detailed study of the metallicity evolution and star formation history of such simulations can assist in predicting LIGO-detectable compact object binary mergers.
Recent simulations of compact binary evolution suggest the compact object merger rate depends
sensitively on the progenitor's metallicity.  Rare low-metallicity star formation during galaxy
assembly can produce more detected compact binaries than typical star formation.   
Using detailed simulations of galaxy and chemical evolution, we determine how sensitively the compact binary populations
of galaxies   with similar
present-day appearance depend on the details of their assembly.
We also demonstrate by concrete example the extent to which dwarf galaxies overabundantly produce compact binary
mergers, particularly binary black holes, relative to more massive galaxies. 
We discuss the implications for transient multimessenger astronomy with compact binary sources.

@@

% POINT: Huge detail possible
The host galaxies of distant short GRBs have already been extensively investigated, with the associations being used to
draw preliminary conclusions about their progenitors  \citep{2014ARAA..52...43B}.   
%Of necessity, these analyses  on bulk correlations, often stacked to increase significance across multiple sources
%\editremark{clean up/don't piss off   Edo}.  
Unlike GRBs, detected gravitational wave sources will be limited by the range of LIGO to the local universe; for
example,  binary neutron star sources should be closer than $400\unit{Mpc}$.  Due to their proximity, each host galaxy
can be explored at great depth and detail via  position-resolved spectroscopy, enabling detailed position-resolved star-formation
and chemical evolution histories  \citep[see,e.g.][]{2009MNRAS.396..462K,2014MNRAS.444..336C,
CALIFA,CALIFA2}.  
%
However, unlike short GRBs and supernovae (SN), present-day compact  binary populations can depend sensitively on rare low metallicity
star formation.
% POINT: But is it useful?
In this work, we assess by concrete example the extent to which detailed analysis of individual galaxies' assembly histories will be essential in
investigating key physical questions about the origin of compact binary mergers.

This paper is organized as follows. 
In \S \ref{sec:sims}, we describe detailed hydrodynamical simulations of  several galaxies, including four of Milky Way-mass and two dwarfs.  Though the four Milky Way-mass galaxies are morphologically similar at $z = 0$, we
show that their star formation and chemical evolution history have subtle but important differences due to their
distinctive merger histories.   To demonstrate
these differences have a practical impact,  in \S \ref{sec:model}, we introduce a simple,
metallicity-dependent  phenomenological model to calculate the present-day rate and mass distribution of compact binary
mergers from  a galaxy's known history.   In \S \ref{sec:results:BBH}, we use this model to investigate  the compact
binary  coalescence rate
dependence on the galaxy's assembly history.   
%% We also demonstrate that even the coalescence rates of binary neutron stars (NSNS) and black hole-neutron
%% star (BH-NS) binaries also depend surprisingly sensitively on the assembly history of each galaxy. 
In \S~\ref{sec:Discussion} we discuss the implications of our study for the interpretation of host galaxy associations
identified via transient multmiessenger astronomy, in the short and long term.  We summarize our results in \S~ \ref{sec:conclude}.
%% In \S \ref{sec:results} we show that while these two galaxy formation models should produce comparable detectable binary neutron star
%% populations, these two scenarios could produce dramatically different populations of electromagnetically detectable
%% merging black hole-neutron star binaries. 
%

@@

\subsection{Simulating galaxy evolution}
To thoroughly study the significance of low metallicity star
formation, we examine cosmological smoothed particle hydrodynamics
(SPH) $N$-body simulations of Milky Way-like galaxies with GASOLINE
\citep{Stadel01,Wadsley04}.  These simulations allow us to analyze both
spatially and temporally resolved star formation, and determine the metallicity history of compact object progenitors.%, but also their dynamical evolution.  
%maybe move the below section to later and/or consolidate
%One simulation we report here, h258,
%involves the formation of a Milky Way-like disc galaxy whose
%progenitors undergo major mergers at $z = 2$ and $z = 0.8$.  This
%simulation has been previously discussed by
%\citet{Governato09,Bellovary10,Bellovary11,Brooks11}; however we have
%re-run the simulation at a factor of two higher in spatial resolution
%and additionally included new physics including a prescription for
%cooling by molecular hydrogen \citet{Christensen12}.

We selected our simulated regions of interest from a volume of
uniform resolution, and resampled the region at very high
resolution using the volume renormalization technique \citep{Katz93}.
This technique allows us to follow the detailed physical processes
involved in galaxy evolution in our selected region while still
including large-scale torques from cosmic structure.  Our cosmological parameters are $\Omega_0 = 0.24$, $\Omega_{\rm baryon} = 0.04$, $\Lambda = 0.76$, h = 0.73, $\sigma_8 = 0.77$ 
 \citep{WMAP3} and we model the ionizing UV background
with the prescription from \citet{Haardt96}.  Our interstellar medium (ISM) model includes the non-equilibrium formation and destruction of H$_2$, which is incorporated in the cooling model along with metal lines \citep{Christensen12}.  Stars form
probabilistically from gas particles which meet density ($n_{min} =
0.1$ amu cm$^{-3}$) and temperature ($T_{max} = 10^3$ K) thresholds, though since star formation also depends on the H$_{\rm 2}$ content of a particle (see below) the densities are nearly always much higher than this threshold. If a gas
particle meets these criteria, it has a likelihood of forming a star
particle (representing a simple stellar population with a Kroupa IMF
\citep{Kroupa93}) which is given by

@@

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{Kennicutt89};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses; \footnote{Gas particles start with a set mass
 and may gain mass from feedback and lose it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{Christensen10} for a discussion
of resolution issues in SPH simulations.} 
 $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  A detailed study of different ISM models and the resulting star formation properties by \citet{Christensen14a} demonstrates that this model allows star formation to occur in clumps of dense gas, comparable to giant molecular clouds.

We model supernova feedback using the
blastwave formalism described in \citet{McKee77} and implemented in
our simulations as in \citet{Stinson06}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
adiabatic expansion phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{Brooks07}, % JB: removed Kirby citation
 the Tully-Fisher relation \citep{Christensen16}, and
the size-luminosity relation \citep{Brooks11}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{Governato10,Governato12}, low-mass disk galaxies with bulges \citep{Christensen14b}, and the Milky Way \citep{Guedes11}.

@@

We identify individual galaxies using the halo finder $AHF$
\citep{Gill04,Knollmann09}, which identifies haloes based on an
overdensity criterion for a flat universe \citep{Gross97}.  In each simulation, we are focusing on the stars which make up the primary (i.e. most massive) galaxy within the zoomed-in high resolution region at $z = 0$.%, which serves as the progenitor to our $z = 0$ Milky
%Way-like galaxy.

@@

%{\bf Richard go ahead and edit this however is needed.}  
The evolution
of a galaxy, in terms of its stellar mass and metallicity evolution,
depends strongly on its interaction history.  Galaxies which appear
similar at the present day may have had drastically different
histories, which may result in differences in compact object merger
rates.  To investigate whether galaxy history affects the compact binary event
rate, we have chosen four simulations which are morphologically similar at $z
= 0$  (see Figure \ref{fig:images}) but differ strongly in their merger histories.

The simulation h277  is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{Loebman12,Loebman14,Kassin14}, baryon fraction
\citep{Munshi13}, and satellite properties \citep{Zolotov12,Brooks14}.  It has a virial mass of $M_{vir} =
6.79 \times 10^{11} \msun$, stellar mass $M_* = 4.24 \times 10^{10}
\msun$, and maximum circular velocity $v_{\rm circ} = 235$ km s$^{-1}$.

@@

We include two additional Milky Way simulations with similar $z = 0$ properties, which have evolutionary histories that fall in between the extremes of the two described above.  The galaxy h239 has a total virial mass of $M_{vir} = 9.3 \times 10^{11}\msun$, stellar mass $M_* = 4.50 \times 10^{10}
\msun$, and maximum circular velocity $v_{\rm circ} = 250$ km s$^{-1}$.  The galaxy h285 has a total virial mass of $M_{vir} = 8.82 \times 10^{11}\msun$, stellar mass $M_* = 4.56 \times 10^{10} \msun$, and maximum circular velocity $v_{\rm circ} = 248$ km s$^{-1}$.  These galaxies are also described in \citet{Bellovary14,Sloane16}.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{boring}
\includegraphics[width=\columnwidth]{exciting}
\caption{\label{fig:GalaxyImages}Synthetic SDSS $gri$ images of two of our Milky Way-type galaxies, h277 (left) and h258 (right), created with \textsc{SUNRISE} \citep{Jonsson06}.
\label{fig:images}
}
\end{figure*}

\begin{figure*}
\includegraphics[width=\columnwidth]{sfh}
\includegraphics[width=\columnwidth]{zvst}
\caption{\label{fig:TwoGalaxies}\textbf{Star formation and metallicity versus time}:  \emph{Left panel}: Star formation history $\dot{M}_{*}$
  versus time.  Black corresponds to \BoringGalaxy{} and red to \ExcitingGalaxy.  
\emph{Right panel}: A plot of the  metallicity $Z$ of recently-formed stars versus time.  Solid red and black lines show
the mean metallicity; dotted lines correspond to  90\% of newly-born stars with lower metallicity,   or 10\% of
newly-born stars with greater metallicity.  When these two galaxies are $2-3\unit{Gyr}$ old, prior to the first major
merger, newly-born stars are created with significantly different metallicity.  Additionally, prior to the
second major merger of h258 at $\simeq 6\unit{Gyr}$, stars form in the \ExcitingGalaxy{} galaxy at a systematically lower
metallicity than in the \BoringGalaxy{} counterpart.
}
\end{figure*}

@@

We have also employed the results of detailed simulations for two dwarf galaxies: \DwarfOne{} and \DwarfTwo{}.  The simulation of h603 consists of a low-mass disc galaxy (qualitatively similar to M33).  It has a virial mass of $3.4 \times 10^{11} \msun$, stellar mass of $7.8 \times 10^9 \msun$, and maximum circular velocity of 111 km s$^{-1}$.  The structure and star formation of this galaxy has been extensively studied by \citet{Christensen14b}.  We also include a bulgeless dwarf galaxy, h516, which has a disc with irregularly distributed star formation, with a virial mass of $3.8 \times 10^{10}\msun$, stellar mass $2.5 \times 10^8 \msun$, and maximum circular velocity of 65 km s$^{-1}$.  Images of these galaxies are shown in Figure \ref{fig:dwarfimages}, and we show their star formation history and metallicity evolution in  Figure \ref{fig:TwoDwarfGalaxies}.  Note that these galaxies have quite different masses, and are not meant to be directly comparable.  The more massive h603 has a much more active star formation history and an overall increasing metallicity with time, whereas the less massive h516 is characterized by small bursts of star formation and a fairly flat metallicity evolution.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{dwarf1}
\includegraphics[width=\columnwidth]{h516}
\caption{\label{fig:GalaxyImages}Synthetic SDSS $gri$ images of our two lower-mass galaxies, h603 (left) and h516, created with \textsc{SUNRISE} \citep{Jonsson06}.
\label{fig:dwarfimages}
}
\end{figure*}

\begin{figure*}
\includegraphics[width=\columnwidth]{sfh_dwarf}
\includegraphics[width=\columnwidth]{dwarf_zvst}
\caption{\label{fig:TwoDwarfGalaxies}\textbf{Star formation and metallicity versus time}:  \emph{Left panel}: Star formation history $\dot{M}_{*}$
  versus time.  Red corresponds to h603 and black to h516.  
\emph{Right panel}: A plot of the  metallicity $Z$ of recently-formed stars versus time.  Solid red and black lines show
the mean metallicity; dotted lines correspond to  90\% of newly-born stars with lower metallicity,   or 10\% of
newly-born stars with greater metallicity.  %When these two galaxies are $2-3\unit{Gyr}$ old, prior to the first major
%merger of  near the peak of their
%star-forming history, newly-born stars are created with significantly different metallicity.  Additionally, prior to the
%second major merger of galaxy B at $\simeq 6\unit{Gyr}$, stars form in the \ExcitingGalaxy{} galaxy at a systematically lower
%metallicity than in the \BoringGalaxy{} counterpart.
}
\end{figure*}

@@

\abbrvPSgrb{} and \abbrvPSellipticals{} show this distribution is a reasonable approximation to compact binary delay

@@

The close agreement between the two galaxies' binary black hole populations, determined by the anticorrelation between
star formation rate and metallicity, may be a single example of a broad trend.  This anticorrelation could cause
galaxies with similar present-day properties to always have similar present-day binary black hole populations,
regardless of their detailed assembly histories.
However, it is also possible that this close agreement is entirely a numerical coincidence between the specific star formation rate and metallicities in these two galaxies.  The effects described here
  are well within the scatter around the mass-metallicity relation.  Marginally different realizations of these
  histories can easily produce factors of order unity difference in galaxies with otherwise indistinguishable
  present-day properties.  That said, two other Milky-Way-like  galaxies (h239 and h285) also show very
  similar binary black hole coalescence rates $r_D/M$; see Table \ref{tab:Results}. % \editremark{but: do more examples show dramatic differences, or comparable changes}

@@

\begin{table}
\begin{centering}
\begin{tabular}{llllc}\hline
\text{Simulation}  &  \text{BHBH} &  \text{BHNS}  & \text{NSNS} & M$_* (\msun)$\\
\hline
\hline
 \text{\BoringGalaxy} & 0.0224 & 0.000247 & 0.000206  & $4.24 \times 10^{10}$ \\
 \text{\ExcitingGalaxy}  &0.0216 & 0.000297 & 0.000228 & $4.46 \times 10^{10}$ \\
  \text{h239}  &  0.023 & 0.000354 & 0.000268 & $4.50 \times 10^{10}$\\
 \text{h285} &  0.0236 & 0.000284 & 0.000241 & $4.56 \times 10^{10}$ \\ \hline
 \text{\DwarfOne} &  0.0381 & 0.000308 & 0.000441& $7.8 \times 10^9$ \\
 \text{\DwarfTwo}  & 0.0949 & 0.000257 & 0.000884& $2.5 \times 10^8$ \\
\hline
\end{tabular}
\end{centering}
\caption{\label{tab:Results}Event rates per unit mass $r_D/M$, arbitrary units, for the three different types of binaries discussed, along with the stellar mass of each galaxy.  
}
\end{table}

@@

While a gravitational wave detection provides detailed information about merging objects (i.e. masses, spins, distance), we need further knowledge to understand the actual origins of the progenitors.  Even when electromagnetic counterparts are available, the long delay times between binary formation and merger limit the prospects of examining the environment where the binary first formed.  On the one hand, compact binaries are kicked by supernovae, moving substantially away from their birth position
\citep{2013ApJ...776...18F,2014ARAA..52...43B}. 
On the other hand, particularly during the early epoch of galaxy formation, gas in galaxies is well-mixed: stars and adjacent gas generally do not have similar chemical composition.  These mixing effects have been previously recognized as an obstacle to interpreting transient event spectra.  For
example,   \citet{2010MNRAS.402.1523P} previously demonstrated that absorbing gas neighbouring transient events (there, long GRBs)
would generally have high metallicity, even for low-metallicity progenitors.   
%
\citet{2010MNRAS.402.1523P} have previously used hydrodynamical simulations to demonstrate that observed ambient
metallicities (there, using damped Lyman $\alpha$ absorbers in the host) do not tightly constrain the metallicity
distribution of the progenitor; see, e.g., their Fig 3.  Thus, the metallicity of stars and gas adjacent to a specific merger event provides
few direct, unambiguous clues to a compact binary merger's progenitors.
